---
title: "COVID-19 Germany introductions"
subtitle: "Extract Germany transmission lineages (BEAST + DTA MCC trees)"
author: "BRICS"
date: '`r format(Sys.time(), "Last modified: %d %b %Y")`'
output: 
  github_document:
    toc: true
    toc_depth: 2
layout: page
editor_options: 
  chunk_output_type: inline
params: 
  inputpath  :  "../../../data/phylogenetic/"
  cluster_f  : DTA_MCC
  minsize    : 100
  uklimit    : "2020-01-23"
  cutoffs    : 0.975, 0.95, 0.75, 0.5, 0.25, 0.05, 0.025
  metadata   : "../results/gisaid-20210602-metadata-sampled-unsampled.tsv"
  intreepath : "../results/beast/run/tree-rich/"
  outputpath : "../results/beast/run/lin-rich/"
  state      : "Germany"
  sampledMetadataFile : "../results/gisaid-20210602-metadata-sampled.tsv"
  
---

```{r rsetup, include=FALSE}
    
  # intreepath : "../results/beast/run/tree-rich/"
  # outputpath : "../results/beast/run/lin-rich/"
    library(tictoc)
    library(dplyr)
    library(lubridate)
    library(ape)
    library(tidytree)
    library(parallel)
    library(beastio)
    source("../reports/treeutils.R")
    source("../reports/clusterutils.R")

    inputpath  <- params$inputpath
    outputpath <- params$outputpath
    intreepath <- params$intreepath
    cluster_f  <- params$cluster_f
    minsize    <- params$minsize
    state      <- params$state
  
    uklimit    <- decimal_date(as.Date(params$uklimit))  
    cutoffs    <- as.numeric(unlist(strsplit(params$cutoffs, split=',')))
    cutoffs <- c(cutoffs, NA)
    #DEBUG
    cutoffs <- c(NA) # !!!!!!!

    figpath    <- paste0(outputpath, "figures/")
    cachepath  <- paste0(outputpath, "cacheMCC/")
    dir.create(outputpath, recursive = TRUE, showWarnings = FALSE)
    dir.create(figpath, recursive = TRUE, showWarnings = FALSE)

    knitr::opts_chunk$set(tidy=FALSE, cache=FALSE, cache.path = cachepath, 
                          dev="png", dpi=150, fig.path=figpath, 
                          message=FALSE, error=FALSE, warning=TRUE, echo=FALSE)

    #metadata              <- read.csv(paste0(inputpath, "metadata.csv"))
    metadata <- read.table(params$metadata, sep="\t", head=TRUE, na.strings=c("NA", ""), fill=TRUE, stringsAsFactors=FALSE, quote="|")
    #metadata <- read.table(params$metadata, sep=",", head=TRUE, na.strings=c("NA", ""), fill=TRUE, stringsAsFactors=FALSE, quote="|")
    #metadata$sample_date  <- ymd(metadata$sample_date)
    #metadata$sample_date  <- ymd(metadata$Submission.date)
    #metadata$sample_date  <- ymd(metadata$collection_date)
    metadata$sample_date  <- ymd(metadata$Collection.date)
    #metadata$sample_date  <- ymd(metadata$date_corrected)
    metadata$decimal_date <- decimal_date(metadata$sample_date)    
    #metadata$taxon_label  <- metadata$sequence_name
    metadata$taxon_label  <- metadata$Accession.ID
    #metadata$taxon_label  <- metadata$accession_id
    #metadata$taxon_label <- gsub("/", "_", as.character(metadata$sequence_name))
    metadata <- metadata[, -which(colnames(metadata) == "AA.Substitutions")]
    
    sampled_metadata <- read.table(params$sampledMetadataFile, sep="\t", head=TRUE, na.strings=c("NA", ""), fill=TRUE, stringsAsFactors=FALSE, quote="|")
    metadata$is_sampled <- metadata$Accession.ID %in% sampled_metadata$Accession.ID
    
    treefiles <- list.files(path = intreepath, pattern = "*.tree$")
    
    # #DEBUG
    # treefiles <- c("A-DTA-20210602.MCC.tree", "B.1.221-DTA-20210602.MCC.tree")
    # lineageLimit = sapply(treefiles, function (x) { if (length(grep("^B.1.1.7", x)) > 0) return( decimal_date(as.Date("2020-12-20"))); return(uklimit); } )
    lineageLimit = sapply(treefiles, function (x) { return(uklimit); } )

    
```


# Summary

This notebook extracts TMRCAs and other summary statistics of the `r cluster_f` clusters from MCC trees from BEAST.

## Input
- Metadata table in `.csv` format. Should contain a `sequence_name` and `sample_date` column.
- Set of MCC trees for with DTA reconstructions save in `r intreepath`.

## Steps
- Extract a table of nodes and heights for each lineage in each tree. 
- Extract TMRCAs and sizes for each lineage in each tree.
- Extract subtrees for all lineages with at least `r minsize` tips.
- Extract lineage assignment for all Germany sequences in each tree replicate.
- Extract the lineage assignment for a threshold of 0.5

## Output
- `.csv` files with nodes and heights for each lineage in each tree.
- `.csv` file with lineage assignments for all Germany sequences for each threshold. 
- `.csv` file with lineage assignments for all Germany sequences for threshold of 0.5. 
- Newick trees for all lineages with at least `r minsize` tips. 



# Extract clusters

```{r extractClustersMCC, results="asis", warning=TRUE, message=TRUE, cache=FALSE, eval=TRUE}
       
      
    #tic(cat(paste0("\n\nTotal time for ", replicates, " trees: ", replicates)))
    clusterStats <- c()   
    clusterStatsSingles <- c()   
    firstRound = TRUE
    for (treefile in treefiles) {   
        tic(cat(paste0("\n\n**Loading ", treefile, ":** ")))
        tree       <- read.beast(paste0(intreepath, treefile))
        treemeta   <- droplevels(metadata[match(tree@phylo$tip.label, metadata$taxon_label), ])
        mostrecent <- max(treemeta$decimal_date, na.rm = TRUE)
        write.table(treemeta, paste0(outputpath, gsub("\\.tree", "", treefile), ".metadata.tsv"), row.names=FALSE, quote=FALSE, sep="\t")
        toc()
        treeLimit <- lineageLimit[treefile]

    
        cat(paste0("\n\n - Most recent tip: ", format.Date(round_date(date_decimal(mostrecent), unit = "day")), " (", mostrecent, ")\n"))
        cat(paste0(" - Maximum Germany node height: ", mostrecent - treeLimit, "\n\n"))

        tic(cat("\n - Extract clusters from trees: "))
        treepath <- paste0(outputpath, gsub("\\.tree", "", treefile), "/")
        dir.create(treepath, recursive = TRUE, showWarnings = FALSE)
        
        treeClusterStats <- c()
        treeClusterStatsSingles <- c()
        for (cutoff in cutoffs) {
          
            # Extract clusters
            #prefix    <- paste0(strsplit(treefile, "-")[[1]][1], "_",cluster_f,"_")
            #prefix    <- paste0(gsub(".tree", "", treefile), "_",cluster_f,"_")
	          prefix    <- paste0("LIN-", state, "-", gsub("-DTA-", "-", gsub(".MCC.tree$", "", treefile)), "_", cluster_f, "_")
            #treetable <- getClusterTable(tree, "Germany", prefix = prefix, cutoff = cutoff, maxheight = (mostrecent - uklimit), debug = FALSE) 
            treetable <- getClusterTable(tree, state, prefix = prefix, cutoff = cutoff, maxheight = (mostrecent - treeLimit), debug = FALSE) 
            write.csv(treetable, paste0(treepath, gsub("\\.tree", "_", treefile), cutoff, ".treetable.csv"), quote=FALSE, row.names=TRUE)
            
            
            
            # Should remove singletons?
            clusterSamples      <- treetable[treetable$types == "sample", ]
            #clusterSamples      <- treetable[treetable$types == "sample" & !grepl("^inode[0-9]+$", treetable$taxon), ]
            clusterMeta         <- droplevels(metadata[match(clusterSamples$taxon, metadata$taxon_label), ])
            clusterMeta$cluster <- clusterSamples$cluster[match(clusterMeta$taxon_label, clusterSamples$taxon)]
            
            # Save summary file
            clusterSizes   <- table(clusterMeta$cluster)
            clusterNames   <- names(clusterSizes)[clusterSizes > 1]
            clusterSingles <- names(clusterSizes)[clusterSizes == 1]
            
            mostrecent <- max(clusterMeta$decimal_date, na.rm = TRUE)
            
            tmrcas      <- sapply(clusterNames, function(x) mostrecent - max(treetable$heights[treetable$cluster == x], na.rm = TRUE))
            tmrcaNodes  <- sapply(clusterNames, function(x) which(treetable$heights == max(treetable$heights[treetable$cluster == x], na.rm=TRUE) ))
            heightRange <- t(sapply(seq_along(tmrcas), function(x) {
                                    result <- mostrecent - c(unlist(tree@data$height_0.95_HPD[tree@data$node == tmrcaNodes[x]]), 
                                                             unlist(tree@data$height_range[tree@data$node == tmrcaNodes[x]]))
                                    if (length(result) < 4) {
                                        return( rep(tmrcas[x], 4))
                                    } else {
                                        return(result)
                                    }
                                  }))
            if (length(tmrcas) == 0) {
              heightRange <- data.frame(matrix(ncol=4, nrow=0))
            }
            colnames(heightRange) <- c("tmrca_HPD_upper", "tmrca_HPD_lower", "tmrca_range_upper", "tmrca_range_lower")

            oldest  <- sapply(clusterNames, function(x) min(clusterMeta$decimal_date[clusterMeta$cluster == x], na.rm = TRUE))
            newest  <- sapply(clusterNames, function(x) max(clusterMeta$decimal_date[clusterMeta$cluster == x], na.rm = TRUE))
            seqs    <- sapply(clusterNames, function(x) sum(clusterMeta$cluster == x, na.rm = TRUE))

            
            if (length(tmrcas) == 0) {
              result <- data.frame(matrix(ncol=8, nrow=0))
              colnames(result) <- c("cluster", "seqs", "tmrca", "tmrca_calendar", 
                                 "oldest", "mostrecent", "treefile", "cutoff")
            } else {
              result <- data.frame(cluster=clusterNames, seqs=seqs, tmrca=tmrcas, tmrca_calendar=round_date(date_decimal(tmrcas), unit="day"), 
                                 oldest=oldest, mostrecent=newest, treefile=treefile, cutoff=cutoff)
            }
            result <- cbind(result, heightRange)
            treeClusterStats <- rbind(treeClusterStats, result)
            
            if (length(clusterSingles) == 0) {
              resultSingles <- data.frame(matrix(ncol=5, nrow=0))
              colnames(resultSingles) <- c("cluster", "seqs", "tmrca", "treefile", "cutoff")
            } else {
                        resultSingles <- data.frame(cluster=clusterSingles, 
                                                    seqs=sapply(clusterSingles, function(x) sum(clusterMeta$cluster == x, na.rm = TRUE)), 
                                                    tmrca=sapply(clusterSingles, function(x) mostrecent - max(treetable$heights[treetable$cluster == x], na.rm = TRUE)),
                                                    treefile=treefile, cutoff=cutoff)
              
            }
            treeClusterStatsSingles <- rbind(treeClusterStatsSingles, resultSingles)
            
            # Extract and save trees for all clusters with >minsize tips
            clusterpath <- paste0(treepath, gsub("\\.tree", "_", treefile), cutoff, "/")
            dir.create(clusterpath, recursive = TRUE, showWarnings = FALSE)
            
            bigClusters <- names(seqs)[which(seqs >= minsize)]
            for (cluster in bigClusters) {
                clusterIds  <- as.character(na.omit(clusterSamples$taxon[clusterSamples$cluster == cluster,drop=TRUE]))
                clusterMRCA <- getMRCA(tree@phylo, clusterIds)
                
                #clustertable <- treetable[treetable$cluster == cluster & !is.na(treetable$cluster), ]
                #clustertable <- clustertable[order(clustertable$heights, decreasing=TRUE), ]
                #clusterMRCA2 <- as.numeric(rownames(clustertable)[1])

                clusterTree  <- treeio::tree_subset(tree, clusterMRCA, levels_back=0)    
                #write.tree(clusterTree@phylo, paste0(clusterpath, cluster, ".tree"))
                write.beast(clusterTree, paste0(clusterpath, cluster, ".tree"))
            }
        }
        
        write.csv(treeClusterStats, paste0(outputpath, gsub("\\.tree", "_", treefile), cluster_f,".csv"), row.names=FALSE, quote=FALSE)
        #clusterStats <- rbind(clusterStats, treeClusterStats)
        if (firstRound) {
          write.table(treeClusterStats, paste0(outputpath, "clusters_",cluster_f,".tsv"), row.names=FALSE, quote=FALSE, sep="\t")
        } else {
          write.table(treeClusterStats, paste0(outputpath, "clusters_",cluster_f,".tsv"), row.names=FALSE, quote=FALSE, sep="\t", col.names=FALSE, append = TRUE)
        }

        #clusterStatsSingles <- rbind(clusterStatsSingles, treeClusterStatsSingles)
	      if (firstRound) {
      	   write.table(treeClusterStatsSingles, paste0(outputpath, "clusters_",cluster_f,"_singles.tsv"), row.names=FALSE, quote=FALSE, sep="\t")
      	} else {
      	   write.table(treeClusterStatsSingles, paste0(outputpath, "clusters_",cluster_f,"_singles.tsv"), row.names=FALSE, quote=FALSE, sep="\t", col.names=FALSE, append = TRUE)
      	}
        
        toc()
      	firstRound = FALSE
    }
    # write.csv(clusterStats, paste0(outputpath, "clusters_",cluster_f,".csv"), row.names=FALSE, quote=FALSE)
    #write.table(clusterStats, paste0(outputpath, "clusters_",cluster_f,".tsv"), row.names=FALSE, quote=FALSE, sep="\t")
    #toc()

    #write.table(clusterStatsSingles, paste0(outputpath, "clusters_",cluster_f,"_singles.tsv"), row.names=FALSE, quote=FALSE, sep="\t")
   
    
```

# Extract cluster samples

```{r extractClusterSamples, results="asis", eval=TRUE}
      
    treedirs <- gsub("\\.tree", "", treefiles)
          
    allSamples <- c()     
    firstRound = TRUE
    for (dir in treedirs) {
        tic(cat(paste0("\n\nProcessing ", dir, ": ")))
      
        treesamples <- NULL
        #treetables  <- list.files(path = paste0(outputpath, dir, "/"), pattern = "*.treetable.csv$")
        for (cutoff in cutoffs) {
            treetable        <- read.csv(paste0(outputpath, dir, "/", dir, "_", cutoff, ".treetable.csv"))
            ukseqs           <- droplevels(treetable[treetable$location == "Germany" & treetable$types == "sample", c("taxon", "cluster")])
            colnames(ukseqs) <- c("taxon", paste0("cutoff_", cutoff))
            
            if (is.null(treesamples)) {
                treesamples <- ukseqs[order(ukseqs$taxon), ]
            } else {
                # Reorder
                ukseqs <- ukseqs[match(treesamples$taxon, ukseqs$taxon), ]
                            
                # Check and bind
                if (all(ukseqs$taxon == treesamples$taxon)) {
                    treesamples[[paste0("cutoff_",cutoff)]] <- ukseqs[[paste0("cutoff_",cutoff)]]
                } else {
                   stop(sprintf("Tree with cutoff %d contains different Germany taxa", cutoff))
                }
                
            }
        }
        # allSamples <- rbind(allSamples, treesamples)
        toc()
        if (firstRound) {
          write.table(treesamples, paste0(outputpath, "clusterSamples_",cluster_f,".tsv"), quote=FALSE, row.names=FALSE, sep="\t")
        } else {
          write.table(treesamples, paste0(outputpath, "clusterSamples_",cluster_f,".tsv"), quote=FALSE, row.names=FALSE, sep="\t",  col.names=FALSE, append = TRUE)
        }
      	firstRound = FALSE
    }
    # write.csv(allSamples, paste0(outputpath, "clusterSamples_",cluster_f,".csv"), quote=FALSE, row.names=FALSE)
    # write.table(allSamples, paste0(outputpath, "clusterSamples_",cluster_f,".tsv"), quote=FALSE, row.names=FALSE, sep="\t")


```

# Extract 50%, 75% and 95% clusterings

```{r extractThreshold, results="asis"}
      
    thresholds <- c(0.5, 0.75, 0.95, NA)   
      
    # clusterStatsAll <- read.csv(paste0(outputpath, "clusters_",cluster_f,".csv"))
    # clusterSamples  <- read.csv(paste0(outputpath, "clusterSamples_",cluster_f,".csv"))
    clusterStatsAll <- read.table(paste0(outputpath, "clusters_",cluster_f,".tsv"), sep="\t", head=TRUE, na.strings=c("NA", ""), fill=TRUE, stringsAsFactors=FALSE, quote="|")
    clusterSamples  <- read.table(paste0(outputpath, "clusterSamples_",cluster_f,".tsv"), sep="\t", head=TRUE, na.strings=c("NA", ""), fill=TRUE, stringsAsFactors=FALSE, quote="|")
    
    cat("\n\n")
    for (threshold in thresholds) {
        if (is.na(threshold)) {
          clusterStats <- clusterStatsAll[is.na(clusterStatsAll$cutoff), ]
      	} else {
              clusterStats <- clusterStatsAll[clusterStatsAll$cutoff == threshold, ]
      	}
        # write.csv(clusterStats, paste0(outputpath, "clusters_",cluster_f,"_", threshold, ".csv"), quote=FALSE, row.names=FALSE)
        write.table(clusterStats, paste0(outputpath, "clusters_",cluster_f,"_", threshold, ".tsv"), quote=FALSE, row.names=FALSE, sep="\t")
        
        clusterMeta         <- droplevels(metadata[match(clusterSamples$taxon, metadata$taxon_label), ])
        clusterMeta$cluster <- clusterSamples[match(clusterMeta$taxon_label, clusterSamples$taxon), paste0("cutoff_", threshold)]
        
        # write.csv(clusterMeta, paste0(outputpath, "clusterSamples_",cluster_f,"_", threshold,".csv"), quote=FALSE, row.names=FALSE)
        write.table(clusterMeta, paste0(outputpath, "clusterSamples_",cluster_f,"_", threshold,".tsv"), quote=FALSE, row.names=FALSE, sep="\t")
        
        cat(paste0("- Posterior probability limit ", threshold, " (", nrow(clusterStats), " transmission lineages)\n"))  
    }

    cat("\n\n")
```




# Session info

```{r sessionInfo, results='markup'}
    sessionInfo()
```
